Group assignment

### 2022-11-05 Draft.Rmd
Update Draft.Rmd

df2$country <- as.factor(df2$addresscountryname)
df2 <- select(df2, -addresscountryname)
```
* The original *accommodationtype* variable was transformed into new *acctype_f* factor variable.
```{r}
df2 <- separate(df2, accommodationtype, '@', into = 
                 c('word', 'acctype'))
df2 <- select(df2, -word)
df2 <- mutate(df2, acctype_f = factor(acctype))
```
* *trueprice* variable was generated by dividing the original *price* variable by *price_night*
```{r}
df2 <- separate(df2, price_night, ' ', into =
                     c('pr','word', 'nights', 'night'))
df2 <- select(df2, -pr)
df2 <- select(df2, -night)
df2 <- select(df2, -word)
df2$nights <- as.numeric(df2$nights)
df2$trueprice <- df2$price/df2$nights
```
* *season* variable was created based on the *month* variable
```{r}
df2 <- df2 %>%
  mutate(season = case_when(month == 12 | month == 1 |month == 2 ~ 'winter',
                            month == 3 | month == 4 |month == 5 ~ 'spring',
                            month == 6 | month == 7 |month == 8 ~ 'summer',
                            month == 9 | month == 10 |month == 11 ~ 'autumn'))
df2$season <- as.factor(df2$season)
```
* With the help of the existing *starrating* variable new *class* variable was generated
```{r}
df2 <- df2 %>%
  mutate(class = case_when(starrating == 1.0 | starrating == 1.5 ~ 'tourist',
                           starrating == 2.0 | starrating == 2.5 ~ 'standard',
                           starrating == 3.0 | starrating == 3.5 ~ 'comfort',
                           starrating == 4.0 | starrating == 4.5 ~ 'first class',
                           starrating == 5.0 ~ 'luxury'))
df2$class <- as.factor(df2$class)
```

### • Quality of the dataset - missing values overview and frequency distribution of numeric values

@@

### Frequency plot of observations

@@

## • Summary for the main variables in the transformed dataset
## *Numeric variables*
```{r echo=FALSE}
datasummary(distance + starrating + actualrating + trueprice ~ Mean + Min + Max + SD, data = df2) %>%
   kableExtra::kable_styling(latex_options = "hold_position")
```
### There are prominent outliers for *distance* and *true price* variables, therefore, the dataset was transformed, setting the maximum values at the 99% percentile (35 miles and 600 EUR respectively).

```{r echo=FALSE}
df2 <- df2 %>% filter(distance < 35) %>% filter(trueprice < 600)
datasummary(distance + trueprice ~ mean + Min + Max + SD, data = df2) %>%
   kableExtra::kable_styling(latex_options = "hold_position")
```

### • Distance
### 75% of the hotels are located between 0 and 2.4 miles from the main city center. 
### The histogram below shows the distribution of the distances from the city center.

@@

### • Starrating
### Most of hotels from the dataset have 3 and 4 stars.
```{r echo=FALSE}
ggplot(data=df2, aes(x=starrating)) + geom_histogram(fill='navyblue', color='white', bins = 35) + 
  labs(x='Stars', y='Count') + xlim(1,5)
```

### • Actualrating
### As we can see from the histogram below, 80% of the guest reviews are between 3.4 and 4.1.
```{r echo=FALSE, message=FALSE}
ggplot(data=df2, aes(x=actualrating)) + geom_histogram(fill='navyblue', color='white') + 
  labs(x='Guest reviews', y='Count')
```

### • Trueprice
### The distribution of the prices is skewed to the right, therefore, we can conclude that the mean price is greater then the mode.

@@

### • Year and month
### 22270 records from the dataset are for the year 2017 and 31570 - for 2018.
```{r echo=FALSE}
kable(count(df2, year)) %>%
  kableExtra::kable_styling(position = "center", latex_options = "hold_position")
```

### Winter and spring are fully presented in the dataset, whereas there are no records for July, August, September, and October.

```{r echo=FALSE}
kable(count(df2, month)) %>%
  kableExtra::kable_styling(position = "center", latex_options = "hold_position")
```

## *Factor variables*

### • Accommodation type
### There are 16 unique accommodation types in the dataset. The most numerous types are "Hotel",  "Apartment", "Guest House",  and "Bed and Breakfast". Together they make up 93% of all records.

@@

### • Country
### The countries with the most records are Italy and Germany. Together they make up 66% of all records.
```{r echo=FALSE}
kable(count(df2, country) %>% arrange(desc(n))) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

### • Season and class
### As we observed earlier, the dataset did not include two months for summer and two months for autumn, consequently, spring and winter have the most observations.
```{r echo=FALSE}
kable(count(df2, season) %>% arrange(desc(n))) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

@@

### • Weekend and holiday
### 35362 records were made on weekends (66%),  and 18478 - on weekdays (34%)

@@

## 1.3 Analysis 
### For this project, we wanted to find the best deal among different types of accommodation in Central European countries. 
### Consequently, we formulated the following question: 
* *How does the average price for accommodation change for different values of the following variables: year, country, distance, season, class, weekend?*

### • As we can see from the graph below, the price increased for 6 countires, Slovakia being the only case in which the price dropped.
### The biggest jump in price was in Italy. In this country, the price in 2018 increased by 24 euros compared to 2017.

@@

>>>>

df2$country <- as.factor(df2$addresscountryname)
df2 <- select(df2, -addresscountryname)
```
* The original *accommodationtype* variable was transformed into new *acctype_f* factor variable.
```{r}
df2 <- separate(df2, accommodationtype, '@', into = 
                 c('word', 'acctype'))
df2 <- select(df2, -word)
df2 <- mutate(df2, acctype_f = factor(acctype))
```
* *trueprice* variable was generated by dividing the original *price* variable by *price_night*
```{r}
df2 <- separate(df2, price_night, ' ', into =
                     c('pr','word', 'nights', 'night'))
df2 <- select(df2, -pr)
df2 <- select(df2, -night)
df2 <- select(df2, -word)
df2$nights <- as.numeric(df2$nights)
df2$trueprice <- df2$price/df2$nights
```
* *season* variable was created based on the *month* variable
```{r}
df2 <- df2 %>%
  mutate(season = case_when(month == 12 | month == 1 |month == 2 ~ 'winter',
                            month == 3 | month == 4 |month == 5 ~ 'spring',
                            month == 6 | month == 7 |month == 8 ~ 'summer',
                            month == 9 | month == 10 |month == 11 ~ 'autumn'))
df2$season <- as.factor(df2$season)
```
* With the help of the existing *starrating* variable new *class* variable was generated
```{r}
df2 <- df2 %>%
  mutate(class = case_when(starrating == 1.0 | starrating == 1.5 ~ 'tourist',
                           starrating == 2.0 | starrating == 2.5 ~ 'standard',
                           starrating == 3.0 | starrating == 3.5 ~ 'comfort',
                           starrating == 4.0 | starrating == 4.5 ~ 'first class',
                           starrating == 5.0 ~ 'luxury'))
df2$class <- as.factor(df2$class)
```
\newpage
## 1.2  Quality of the dataset
### • Missing values overview and frequency distribution of numeric values

@@

## 1.3 Descriptive statistics
## • Summary for the main variables in the transformed dataset
## *Numeric variables*
```{r echo=FALSE}
datasummary(distance + starrating + actualrating + trueprice ~ Mean + Min + Max + SD, data = df2) %>%
   kableExtra::kable_styling(latex_options = "hold_position")
```
### There are prominent outliers for *distance* and *true price* variables, therefore, the dataset was transformed, setting the maximum values at the 99% percentile (35 miles and 600 EUR respectively).

```{r echo=FALSE}
df2 <- df2 %>% filter(distance < 35) %>% filter(trueprice < 600)
datasummary(distance + trueprice ~ mean + Min + Max + SD, data = df2) %>%
   kableExtra::kable_styling(latex_options = "hold_position")
```
### Frequency plot of observations

@@

### 1.3.1 Distance
### 75% of the hotels are located between 0 and 2.4 miles from the main city center. 
### The histogram below shows the distribution of the distances from the city center.

@@

### 1.3.2 Starrating
### Most of hotels from the dataset have 3 and 4 stars.
```{r echo=FALSE}
ggplot(data=df2, aes(x=starrating)) + geom_histogram(fill='navyblue', color='white', bins = 35) + 
  labs(x='Stars', y='Count') + xlim(1,5)
```

### 1.3.3 Actualrating
### As we can see from the histogram below, 80% of the guest reviews are between 3.4 and 4.1.
```{r echo=FALSE, message=FALSE}
ggplot(data=df2, aes(x=actualrating)) + geom_histogram(fill='navyblue', color='white') + 
  labs(x='Guest reviews', y='Count')
```

### 1.3.4 Trueprice
### The distribution of the prices is skewed to the right, therefore, we can conclude that the mean price is greater then the mode.

@@

### 1.3.5 Year and month
### 22270 records from the dataset are for the year 2017 and 31570 - for 2018.
```{r echo=FALSE}
kable(count(df2, year)) %>%
  kableExtra::kable_styling(position = "center", latex_options = "hold_position")
```
### Winter and spring are fully presented in the dataset, whereas there are no records for July, August, September, and October.

```{r echo=FALSE}
kable(count(df2, month)) %>%
  kableExtra::kable_styling(position = "center", latex_options = "hold_position")
```
## *Factor variables*

### 1.3.6 Accommodation type
### There are 16 unique accommodation types in the dataset. The most numerous types are "Hotel",  "Apartment", "Guest House",  and "Bed and Breakfast". Together they make up 93% of all records.

@@

### 1.3.7 Country
### The countries with the most records are Italy and Germany. Together they make up 66% of all records.
```{r echo=FALSE}
kable(count(df2, country) %>% arrange(desc(n))) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

### 1.3.8 Season and class
### As we observed earlier, the dataset did not include two months for summer and two months for autumn, consequently, spring and winter have the most observations.
```{r echo=FALSE}
kable(count(df2, season) %>% arrange(desc(n))) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

@@

### 1.3.9 Weekend and holiday
### 35362 records were made on weekends (66%),  and 18478 - on weekdays (34%)

@@

@@

## 1.4 Analysis 
### For this project, we wanted to find the best deal among different types of accommodation in Central European countries. 
### Consequently, we formulated the following question: 
* *How does the average price for accommodation change for different values of the following variables: year, country, distance, season, class, weekend?*

## 1.4.1 Time
### • As we can see from the graph below, the price increased for 6 countires, Slovakia being the only case in which the price dropped.
### The biggest jump in price was in Italy. In this country, the price in 2018 increased by 24 euros compared to 2017.

## 1.4.2 Country

# Graph to show the correlation pattern by each country:
```{r echo=FALSE}
ggplot(corr_table1, aes(x = correlation ,
                       y = fct_reorder(country, correlation))) +
  geom_point(color = 'red', size = 2)+
  labs(y='Countries',x='Correlation')
```

## 1.4.3 Distance

### The correlation between truelprice and distance 
```{r echo=FALSE}
corr_table1 <- df2 %>% 
  select(country, trueprice, distance) %>% 
  group_by(country) %>% 
  summarise(correlation = cor(trueprice,distance))
```

```{r echo=FALSE}
kable(corr_table1)
```

#### As we may see, staying in the center is less valued in Austria and Slovakia.

### The correlation between guest reviewing report with distance and price
```{r echo=FALSE}
corr_table2 <- cor(df2[, c('trueprice','actualrating','distance')])
```
```{r echo=FALSE}
kable(corr_table2)
```

## 1.5 Hypothesis testing

### 1.5.1 Is there any statistical significant difference in the hotel price per night between year 2017 and 2018
```{r echo=FALSE}
t.test(df2$trueprice [df2$year=="2017"],df2$trueprice [df2$year=="2018"], mu=0, alt='two.sided', conf=0.95, paired=F)
#(welsch1)
#Null Hypothesis is rejected though the t-statistic is insignificant 
```
### 1.5.2 There is no statistically significant difference in the price of hotels between summer and winter season
```{r echo=FALSE}
t.test(df2$trueprice [df2$season=="winter"],df2$trueprice [df2$season=="summer"], mu=0, alt='two.sided', conf=0.95, paired=F)
#Null Hypothesis is rejected
